On Math

### 2025-04-29 cards/paper.tex
end behavior of polynomials

%TODO: image
\begin{equation}
  f(x) = \frac{1}{x}
\end{equation}
When function is not transformed, it passes through points $(-1, -1)$ and $(1, 1)$.
% TODO: not range?
Domain of (transformed) reciprocal function
\begin{equation}
f(x) = \frac{a}{bx+c} + d
\end{equation}
has the horizontal asymptote $y = d$ therefore the range is
\begin{equation}
f(x) = \in (-\infty, d) \cup (d, \infty)
\end{equation}
% TODO: range
Range of (transformed) reciprocal function
\begin{equation}
f(x) = \frac{a}{bx+c} + d
\end{equation}
has the vertical asymptote $y = d$ therefore the range is
\begin{equation}
f(x) = \in (-\infty, d) \cup (d, \infty)
\end{equation}

@@

%TODO: some heading?

@@

\section{Polynomial identities}
\begin{flalign*}
& (a + b)^2 = a^2 + 2ab + b^2 & \\
& (a - b)^2 = a^2 - 2ab + b^2 & \\
& (a + b)^3 = a^3 + 3a^2b + 3ab^2 + b^3 & \\
& (a - b)^3 = a^3 - 3a^2b + 3ab^2 - b^3 & \\
& a^2 - b^2 = (a + b)(a - b) & \\
& a^3 - b^3 = (a - b)(a^2 + ab + b^2) & \\
& a^3 + b^3 = (a + b)(a^2 - ab + b^2) &
\end{flalign*}

\section{Derivatives}
Average rate of change of the function $f(x)$ on the interval $[a, a+h]$:

>>>>

% TODO: image
\begin{equation}
  f(x) = \frac{1}{x}
\end{equation}
When function is not transformed, it passes through points $(-1, -1)$ and $(1, 1)$.
% TODO: not range?
Domain of (transformed) reciprocal function
\begin{equation}
  f(x) = \frac{a}{bx+c} + d
\end{equation}
has the horizontal asymptote $y = d$ therefore the range is
\begin{equation}
  f(x) = \in (-\infty, d) \cup (d, \infty)
\end{equation}
% TODO: range
Range of (transformed) reciprocal function
\begin{equation}
  f(x) = \frac{a}{bx+c} + d
\end{equation}
has the vertical asymptote $y = d$ therefore the range is
\begin{equation}
  f(x) = \in (-\infty, d) \cup (d, \infty)
\end{equation}

@@

% TODO: some heading?

@@

\section{Polynomials}
\subsection{Polynomial identities}
\begin{flalign*}
  & (a + b)^2 = a^2 + 2ab + b^2 & \\
  & (a - b)^2 = a^2 - 2ab + b^2 & \\
  & (a + b)^3 = a^3 + 3a^2b + 3ab^2 + b^3 & \\
  & (a - b)^3 = a^3 - 3a^2b + 3ab^2 - b^3 & \\
  & a^2 - b^2 = (a + b)(a - b) & \\
  & a^3 - b^3 = (a - b)(a^2 + ab + b^2) & \\
  & a^3 + b^3 = (a + b)(a^2 - ab + b^2) &
\end{flalign*}
\subsection{End behavior of general polynomials}
Given polynomial
\begin{equation}
  f(x) = a_nx^n + a_{n-1}x^{n-1} + ... +a_1x + a_0
\end{equation}
with leading coefficient $a \neq 0$ the end behavior is always
\begin{equation}
  f(x) \rightarrow \infty
\end{equation}
or
\begin{equation}
  f(x) \rightarrow -\infty
\end{equation}
as
\begin{equation}
  x \rightarrow \pm \infty
\end{equation}
If the leading coefficient is positive then
$f(x) \rightarrow \infty$ as $x \rightarrow \infty$,
if negative then
$f(x) \rightarrow -\infty$ as $x \rightarrow \infty$
If leading exponent is \textbf{even}, then the end behavior as $x \rightarrow \infty$ \textbf{is the same} as the end behavior as $x \rightarrow -\infty$.
Otherwise, if the leading exponent is \textbf{odd}, then the end behavior as $x \rightarrow +\infty$ is the opposite as $x \rightarrow -\infty$.
\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|l|l|}
    \hline
    \textbf{$n$} & \textbf{$a_n$} & \textbf{$x \rightarrow -\infty$} & \textbf{$x \rightarrow +\infty$} \\
    \hline
    even & $>0$ & $f(x) \rightarrow +\infty$ & $f(x) \rightarrow +\infty$ \\
    \hline
    even & $<0$ & $f(x) \rightarrow -\infty$ & $f(x) \rightarrow -\infty$ \\
    \hline
    odd & $>0$ & $f(x) \rightarrow -\infty$ & $f(x) \rightarrow +\infty$ \\
    \hline
    odd & $<0$ & $f(x) \rightarrow +\infty$ & $f(x) \rightarrow -\infty$ \\
    \hline
  \end{tabular}
  \caption{End behavior of polynomials}
  \label{tab:end_behavior_of_polynomials}
\end{table}
Mnemonic trick:
substitute $+1$ and $-1$ into the leading term of polynomial, for example:
\begin{equation}
  \begin{gathered}
    P(x) = 2x^3 - 6x^2 + 8 \\
    \\
    2(+1)^3 = +2 \\
    P(x) \rightarrow +\infty \\
    \text{as} \\
    x \rightarrow +\infty \\
    \\
    2(-1)^3 = -2 \\
    P(x) \rightarrow -\infty \\
    \text{as} \\
    x \rightarrow -\infty \\
  \end{gathered}
\end{equation}
In general whenever $n$ is an \textit{even natural number} we have
\begin{equation}
  \begin{gathered}
    \lim_{x \to +\infty}x^n = +\infty \\
    \text{and} \\
    \lim_{x \to -\infty}x^n = +\infty
  \end{gathered}
\end{equation}
for $n = 2, 4, 6, ...$ but for \textit{odd natural number} though:
\begin{equation}
  \begin{gathered}
    \lim_{x \to +\infty}x^n = +\infty \\
    \text{and} \\
    \lim_{x \to -\infty}x^n = -\infty
  \end{gathered}
\end{equation}
for $n = 1, 3, 5, ...$. To compute the limit at infinity of general polynomial, we only need
to consider the term with the largest expontent, called the \textbf{leading term}.
All of the other terms can be disregarded.
\begin{equation}
  \begin{gathered}
    \lim_{x \to \infty}(2x^3 - 7x^2 + 6x + 12) = \\
    = \lim_{x \to \infty}(2x^3) = \\
    = \lim_{x \to \infty}(x^3) = \\
    = \infty
  \end{gathered}
\end{equation}
\section{Derivatives}
Average rate of change of the function $f(x)$ on the interval $[a, a+h]$:
